---
title: "Plots and Tables"
author: "Shravanthi Seshasayee"
date: "4 December 2017"
output: html_document
---
Table 1: Summary -Ethnicity-specific, sex, age, hit, pain, elixhauser, oasis, insurance, MS
Table 2: Unadjusted regression - for three catergories of years, between ethnicity and hit
Table 3: Adjusted regression - overall and for three categories

Plot 1: Unadjusted
Plot 2: Adjusted 

Interactive - trends across time?

```{r}
library(tableone)
library(tidyverse)
library(dplyr)
read.csv("F://SSmimic/Final Code and Write Up/finaltable.csv")
```



# Should marital status and insurance be in the same model as ethnicity? I'd like to remove, they make the results harder to interpret (more variables, harder - e_score can also go )
```{r}
#adjusted global test for all the years

finaltable$insurance<- as.factor(finaltable$insurance)
finaltable$ethnicity<- as.factor(finaltable$ethnicity)
finaltable$marital_status<- as.factor(finaltable$marital_status)
finaltable$ethnicity<- relevel(finaltable$ethnicity, ref = "WHITE")
finaltable$marital_status<- relevel(finaltable$marital_status, ref = "MARRIED")

secondlook <- glm(finaltable$drug ~ finaltable$ethnicity + finaltable$mean.pain + finaltable$gender + finaltable$age +finaltable$oasis, family = binomial(link = "logit"))
summary(secondlook)
confint(secondlook)
#sjp.glm(secondlook)

secondlook$coefficients[2]
secondlook$coefficients[3]
secondlook$coefficients[4]

```


```{r}

# Create labels
boxLabels = c("Black", "Hispanic", "Other")

# Enter summary data. boxOdds are the odds ratios (calculated elsewhere), boxCILow is the lower bound of the CI, boxCIHigh is the upper bound.

df <- data.frame(
  yAxis = length(boxLabels):1,
  boxOdds = c(-0.3515534,-0.002075366, -0.1058666),
  boxCILow = c(-0.431987951,-0.127186838, -0.173518623),
  boxCIHigh = c(-0.2708931821, 0.1244000651, -0.0379375410)
)

# Plot
p <- ggplot(df, aes(x = boxOdds, y = yAxis))
p + geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(aes(xmax = boxCIHigh, xmin = boxCILow), size = .5, height = .2, color = "gray50") +
  geom_point(size = 3.5, color = "orange") +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  scale_y_continuous(breaks = yAxis, labels = boxLabels) +
  scale_x_continuous(breaks = seq(0,7,1) ) +
  coord_trans(x = "log10") +
  ylab("") +
  xlab("Odds ratio (log scale)") +
  annotate(geom = "text", y =1.1, x = 3.5, label ="Model p < 0.001\nPseudo R^2 = 0.10", size = 3.5, hjust = 0) + ggtitle("Intention to remove box turtles from the road")



```



```{r}

exp(coefficients(secondlook))
intercept <- exp(secondlook$coef[1]) # intercept
blk <- exp(secondlook$coef[2])
hisp <- exp(secondlook$coef[3])
othr <- exp(secondlook$coef[4])
pain <- exp(secondlook$coef[5])
male <- exp(secondlook$coef[6])
age <- exp(secondlook$coef[7])
oasis <- exp(secondlook$coef[8])
```

```{r}
X1_range <- seq(from=min(data$X1), to=max(data$X1), by=.01)


```











```{r}
omitted <- select(finaltable, "ethnicity", "formulary_drug_cd", "mean.pain")
omitted <- na.omit(omitted)
View(omitted)
```


```{r}
CreateTableOne(vars=c("gender","age","formulary_drug_cd", group_by = "ethnicity"),data=finaltable)

finaltable %>% group_by(ethnicity) %>% summarise(MeanAge = mean(age), Gender = gender )

unique(finaltable$drug)
```

