---
title: "Cohort"
output: html_document
---

<<<<<<< HEAD
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyverse)
```




```{r}
#Merging insurance and marital status with main table
library(readr)
main_table <- read_csv("~/SSmimic/Cohort/main.table.csv")
msi <- read_csv("~/SSmimic/Medication/msi.csv")
main_table <- merge(main_table, msi, by ="hadm_id")
head(main_table)
```

```{r}
#Merging years with main.table
#>>>>>>> d9d4abefe32dcd4d82d6c5a33251ba031172e8c9
years <- read_csv("~/SSmimic/years.csv")
cohort<- merge(main_table,years,by="hadm_id")
write.csv(cohort, file = "cohort.csv")
```

```{r}

Fentanyl <- read_csv("~/SSmimic/Medication/Fentanyl.csv")
Hydromorphone <- read_csv("~/SSmimic/Medication/Hydromorphone.csv")
Morphine <- read_csv("~/SSmimic/Medication/Morphine.csv")
Oxycodone <- read_csv("~/SSmimic/Medication/Oxycodone.csv")
Percocet <- read_csv("~/SSmimic/Medication/Percocet.csv")
Hydrocodone <- read_csv("~/SSmimic/Medication/Hydrocodone.csv")


#create tables with no duplicated hadm_id

Fent<- Fentanyl %>% distinct(hadm_id, .keep_all = TRUE)
Hydrom<- Hydromorphone %>% distinct(hadm_id, .keep_all = TRUE)
Morph<- Morphine %>% distinct(hadm_id, .keep_all = TRUE)
Oxycod<- Oxycodone %>% distinct(hadm_id, .keep_all = TRUE)
Perco <- Percocet %>% distinct(hadm_id, .keep_all = TRUE)
Hydrocod<- Hydrocodone %>% distinct(hadm_id, .keep_all = TRUE)

#create combined opioid medication table
opioid <- rbind(Fent,Hydrom, Morph, Oxycod, Perco, Hydrocod)
opioid_nohadmid<-  opioid %>% distinct(hadm_id, .keep_all = TRUE)

#create csv
write.csv(opioid, file = "opioid.csv")

#create  table with only IV drug presiscriptions and remove duplicated hadm_id

IVtable<- opioid[(opioid$route == "IV") |(opioid$route == "IV DRIP")|(opioid$route == "IVPCA") , ]
IVtable_nodup<-  IVtable %>% distinct(hadm_id, .keep_all = TRUE)
write.csv(IVtable_nodup,file = "IVtable_nodup.csv")


#create  table with NO IV drug presiscriptions and remove duplicated hadm_id

oralopioid<- opioid[!((opioid$route == "IV") |(opioid$route == "IV DRIP")|(opioid$route == "IVPCA")), ]
oralopioid_nodup<-  oralopioid %>% distinct(hadm_id, .keep_all = TRUE)
write.csv(oralopioid_nodup,file = "oralopioid_nodup.csv")

```

```{r}

IVopioid <- read_csv("~/SSmimic/Medication/IVtable_nodup.csv")
cohort <- read_csv("~/SSmimic/Cohort/cohort.csv")

#merger the tables to ease analysis

finaltable<- left_join(cohort,IVopioid, "hadm_id")

#Replacing NA to 0 and receiving an IV opioid to 1:

finaltable$formulary_drug_cd<- ifelse(is.na(finaltable$formulary_drug_cd),0,1)

#categorize race into 4 categories

finaltable$ethnicity <- ifelse(grepl("BLACK",finaltable$ethnicity),"BLACK", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("WHITE",finaltable$ethnicity),"WHITE", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("HISPANIC",finaltable$ethnicity),"HISPANIC", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("ASIAN",finaltable$ethnicity),"OTHER", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("UN",finaltable$ethnicity),"OTHER", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("MUL",finaltable$ethnicity),"OTHER", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("INDIAN",finaltable$ethnicity),"OTHER", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("SOUTH",finaltable$ethnicity),"OTHER", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("ISLAND",finaltable$ethnicity),"OTHER", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("DECL",finaltable$ethnicity),"OTHER", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("EAST",finaltable$ethnicity),"OTHER", finaltable$ethnicity)
finaltable$ethnicity <- ifelse(grepl("PORTU",finaltable$ethnicity),"HISPANIC", finaltable$ethnicity)

finaltable$ethnicity[is.na(finaltable$ethnicity)]<-"OTHER"

#categorize marital status into 4 categories 

finaltable$marital_status.y <- ifelse(grepl("MARRIED",finaltable$marital_status.y),"MARRIED", finaltable$marital_status.y)
finaltable$marital_status.y <- ifelse(grepl("LIFE PARTNER",finaltable$marital_status.y),"MARRIED", finaltable$marital_status.y)
finaltable$marital_status.y <- ifelse(grepl("DIVORCED",finaltable$marital_status.y),"DIVORCED/WIDOWED", finaltable$marital_status.y)
finaltable$marital_status.y <- ifelse(grepl("SEPARATED",finaltable$marital_status.y),"DIVORCED/WIDOWED", finaltable$marital_status.y)

finaltable$marital_status.y <- ifelse(grepl("WIDOW",finaltable$marital_status.y),"DIVORCED/WIDOWED", finaltable$marital_status.y)
finaltable$marital_status.y <- ifelse(grepl("UN",finaltable$marital_status.y),"OTHER", finaltable$marital_status.y)

finaltable$marital_status.y[is.na(finaltable$marital_status.y)]<-"OTHER"

#categorize insurance into 3 catergories

finaltable$insurance.y <- ifelse(grepl("Medi",finaltable$insurance.y),"Government", finaltable$insurance.y)

#check that ethnicity, marital status and insurance were correctly categorised

unique(finaltable$ethnicity)

unique(finaltable$marital_status.y)

unique(finaltable$insurance.y)

#exclude patients with an Elixhauser score that is 10 or more
finaltable<- filter(finaltable, finaltable$e_score < 10)

#filter out years 2001 and 2012 as they are incomplete years and select for relevant columns

finaltable<- filter(finaltable, !finaltable$admityear == "2001" & !finaltable$admityear == "2012")

#select relevant columns

finaltable<- select(finaltable,"subject_id.x", "hadm_id", "med.pain", "age", "gender", "ethnicity","max_pain", "mean.pain", "oasis", "e_score", "admityear", "drug", "formulary_drug_cd", "drug_name_generic", "marital_status.y","insurance.y")

#write csv final table: cohort plus IV opioids woth no duplicated hadm_id's
write.csv(finaltable, file = "finaltable.csv")

```
